---
title: "doc2vec"
author: "Kohei Watanabe"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{doc2vec}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

In this article, **wordvector** is used to train [doc2vec](https://doi.org/10.48550/arXiv.1405.4053) on **quanteda**'s tokens objects.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, collapse = TRUE)
if (identical(Sys.getenv("NOT_CRAN"), "false"))
    options(quanteda_threads = 2, wordvector_threads = 2)
```

## Prepare data

Download the corpus of news summaries.

```{r, eval=TRUE, cache=TRUE}
# Load data
f <- tempfile()
download.file('https://www.dropbox.com/s/e19kslwhuu9yc2z/yahoo-news.RDS?dl=1', 
              f, mode = "wb")
```

```{r}
library(quanteda)
library(wordvector)
quanteda_options(verbose = TRUE)

# Construct corpus
dat <- readRDS(f)
dat$text <- paste0(dat$head, ". ", dat$body)
corp <- corpus(dat, text_field = 'text')

# Tokenize
toks <- tokens(corp, remove_punct = TRUE, remove_symbols = TRUE) %>% 
    tokens_remove(stopwords("en", "marimo"), padding = TRUE) %>% 
    tokens_select("^[a-zA-Z-]+$", valuetype = "regex", case_insensitive = FALSE,
                  padding = TRUE)
```

## Train doc2vec

`textmodel_doc2vec()` supports both `dm` (distributed memory) and `dbow` (distributed bag-of-words) models.

```{r, cache=TRUE}
# Set the number of processors
options(wordvector_threads = 16)

# Train doc2vec
dov <- textmodel_doc2vec(toks, dim = 50, type = "dm", min_count = 5, verbose = TRUE)
```

Since the distributed memory model has hidden layers for documents and words, you can extract document and word vectors using `as.matrix()`

```{r}
# Extract document vector
dim(as.matrix(dov, layer = "documents"))

# Extract word vector
dim(as.matrix(dov, layer = "words"))
```

## Predict probability

If `probabitliy()` is applied to a fitted doc2vec model, you receive the predicted probability of the words in each document.

```{r}
head(probability(dov, c("bad", "good"), mode = "numeric", layer = "documents"))
```

