---
title: "word2vec"
author: "Kohei Watanabe"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{word2vec}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

This article shows how to fit [word2vec](https://arxiv.org/abs/1310.4546) using **wordvector**. Since the model is trained on **quanteda**'s tokens objects, users can easily integrate the model in their current work flow. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, collapse = TRUE)
```

## Prepare data

Download the corpus of news summaries.

```{r, eval=FALSE}
# Load data
f <- tempfile()
download.file('https://www.dropbox.com/s/e19kslwhuu9yc2z/yahoo-news.RDS?dl=1', 
              f, mode = "wb")
```

```{r echo=FALSE}
f <- "~/yahoo-news.RDS"
```

```{r}
library(quanteda)
library(wordvector)
quanteda_options(verbose = TRUE)

# Construct corpus
dat <- readRDS(f)
dat$text <- paste0(dat$head, ". ", dat$body)
corp <- corpus(dat, text_field = 'text')

# Tokenize
toks <- tokens(corp, remove_punct = TRUE, remove_symbols = TRUE) %>% 
    tokens_remove(stopwords("en", "marimo"), padding = TRUE) %>% 
    tokens_select("^[a-zA-Z-]+$", valuetype = "regex", case_insensitive = FALSE,
                  padding = TRUE)
```

## Train word2vec

`textmodel_word2vec()` supports both `cbow` (continuous bag-of-words) and `sg` (skip-gram) models.

```{r}
# Set the number of processors
options(wordvector_threads = 16)

# Train word2vec
wov <- textmodel_word2vec(toks, dim = 50, type = "cbow", min_count = 5, verbose = TRUE)
```

```{r}
# Extract word vector
dim(as.matrix(wov))
```

## Find similar words

`similarity()` is a user-friendly utility function to find similar words. Using the `analogy()` wrapper, you can also add or subtract word vectors before finding similar words. 

```{r}
head(similarity(wov, "bad"))
head(similarity(wov, analogy(~ good - bad)))
```

## Predict probability

As a language model, the model allows you to compute the probability of the word conditional on context words.

```{r}
head(probability(wov, "bad", mode = "numeric"))
```

